---
title: "p8105_hw3_rx2236"
author: "Ruiqi Xue"
date: "2023-10-11"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
```


## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```

```{r}
instacart = 
  instacart |> 
  as_tibble()
```


This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.


```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```


Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2
Import data
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

Clean the dataset
```{r}
brfss_smart2010 =
  brfss_smart2010 |> 
  as_tibble() |>
  janitor::clean_names() |>
  rename("location_abbr" = "locationabbr") |>
  rename("location_desc" = "locationdesc")

brfss_oh = 
  filter(brfss_smart2010, topic == "Overall Health")
 
```

Check values of `response`
```{r}
print(unique(pull(brfss_oh, response)))
```

Keep cleaning the data
```{r}
brfss_oh = 
  filter(brfss_oh, response == "Excellent" |
           response == "Very good" |
           response == "Good" |
           response == "Fair" |
           response == "Poor") |>
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"),
                         ordered = TRUE))


```

Check which states were observed at 7 or more locations in 2002 and 2010.

```{r}
county_2002 = brfss_oh |>  
  filter(year == 2002) |> 
  group_by(location_abbr, location_desc) |> 
  summarise()
  

county_2002 = county_2002 |> 
  group_by(location_abbr) |> 
  summarise(count = n()) |> 
  filter(count >=7) 

print(county_2002)

```

The states that were observed at 7 or more locations in 2002 are: `r county_2002$location_abbr`, which are 6 states.

```{r}
county_2010 = brfss_oh |>  
  filter(year == 2010) |> 
  group_by(location_abbr, location_desc) |> 
  summarise()
  

county_2010 = county_2010 |> 
  group_by(location_abbr) |> 
  summarise(count = n()) |> 
  filter(count >=7) 

print(county_2010)

```

Similarly, we get states that were observed at 7 or more locations in 2010: `r county_2010$location_abbr`, which are 14 states.

Now make a "spaghetti” plot of average data value over time within a state.
```{r}
excellent_df = brfss_oh |> 
  filter(response == "Excellent") |> 
  select(location_abbr, year, data_value)

excellent_df = excellent_df |> 
  group_by(location_abbr, year) |> 
  summarise(mean_value = mean(data_value)) |> 
  rename(state = location_abbr)

excellent_df |> 
  ggplot(aes(x = year, y = mean_value, color = state)) +
  geom_line() + 
  theme(legend.position = "bottom") +
  labs(
    title = "Average Data Value Over Time for Each State",
    x = "Year",
    y = "Mean Data Value (of BHR)",
    color = "State"
  ) 

```
From the plot, we can see that the mean data value of BHR for each state fluctuates over time, and with a pattern although not quite obvious that at year 2005, 2007 the mean values are relatively low (decreased), and at year 2006 and 2008 the mean values are relatively high (increased).



Now, make the two-panel plot showing for the years 2006 and 2010, the distribution of `data_value` for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
NY_response = 
  brfss_oh |> 
  filter(year == 2006 | year == 2010, location_abbr == "NY") |> 
  select(location_desc, response, data_value, year) 

NY_response |>
  ggplot(aes(x = response, y = data_value, color = location_desc)) + 
  geom_point() +
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  facet_grid(~year) +
  labs(
    title = "Distribution of Data Value for Responses Among Locations in NY State",
    x = "Response",
    y = "Data Value (of behavioral health risk)",
    color = "Location"
  ) 
```


## Problem 3

First, clean the datasets and merge them into one dataset.
```{r}
demographic_df = read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female")) |>
  mutate(
    education = case_match(
      education,
      1 ~ "Less than high school",
      2 ~ "High school equivalent",
      3 ~ "More than high school"
    )) |>
  mutate(education = factor(education, levels = c("Less than high school", "High school equivalent", "More than high school"), ordered = TRUE)) |>
  drop_na() |>
  filter(age >= 21)


accel_df = 
  read_csv("data/nhanes_accel.csv") |>
  janitor::clean_names() |>
  drop_na()

merged_df = left_join(demographic_df, accel_df)

merged_df2 = merged_df |>
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    values_to = "mims_value",
    names_prefix = "min"
  ) 
```

Produce a table for the number of men and women in each education category.

```{r}
merged_df |> 
  group_by(sex, education) |> 
  summarise(n = n()) |>
  pivot_wider(
    names_from = sex,
    values_from = n) |>
  knitr::kable()

```
Through the table we can see in "high school equivalent" category the number of female is significantly less than male while in the other two categories there are slightly more female than male.


Then create a visualization of the age distributions for men and women in each education category.
```{r}
merged_df |>
  select(sex, age, education) |>
  ggplot(aes(x = sex, y = age, color = education)) + 
  geom_boxplot() +
  theme(legend.position = "bottom") +
  labs(
    title = "Age Distribution for Men and Women in Each Education Category",
    x = "Sex",
    y = "Age",
    color = "Education"
  ) 
```

Females in "high school equivalent" category tend to have higher age than males, while females in "more than high school" tend to have lower average age than males.


Aggregate across minutes to create a total activity variable for each participant.
Plot these total activities (y-axis) against age (x-axis).
```{r}
merged_df2 |>
  group_by(seqn, sex, age, education) |>
  summarise(total_activity = sum(mims_value)) |>
  ggplot(aes(x = age, y = total_activity, color = sex)) +
  geom_point(alpha = .7) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "bottom") +
  facet_grid(~education) +
  labs(
    title = "Total Activities Against Age for Female and Male at Each Education Category",
    x = "Age",
    y = "Total Activity of a day",
    color = "Sex"
  ) 
  
```
For all education categories, people tend to have lower total activity of a 24-hour day when getting older.

For "high school equivalent" category, the difference between female and male of total activity is more obvious than the other two categories. 

In "less than high school" category, male tend to have higher total activity than female, while in the other two categories this relationship is reversed.

In "more than high school" category, both female and male tend to have lower total activity at their early life compared to the other two education categories.


Now make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. 

```{r}
merged_df2 |>
  mutate(minute = as.numeric(minute)) |>
  group_by(minute, sex, education) |>
  summarise(mean_activity = mean(mims_value)) |>
  ggplot(aes(x = minute, y = mean_activity, color = sex)) +
  geom_point(alpha = .1) +
  geom_smooth(se = FALSE) +
  scale_x_continuous(breaks = c(0, 480, 960, 1440)) +
  theme(legend.position = "bottom") +
  facet_grid(~education) +
  labs(
    title = "24-hour Activity Time Courses for Each Education Level",
    x = "Minutes of a day",
    y = "Mean activity over a day",
    color = "Sex"
  ) 

```
Based on the graph, we can see that for all three education levels, female tend to have higher/more activity over a day than male, while this difference is most obvious during the half of the day, i.e., around noon, and around midnight the activity level is pretty similar for female and male.

It is worth noting that for people in "less than high school" education level group, male's activity level is higher than female around midnight/at nighttime, which is different than the other two groups.

One pattern is that people have higher education level tend to have lower activity level during the daytime, which implies they tend to work in door or their work require less physical activity. And male performs more obviously in this pattern. 








