---
title: "p8105_hw3_rx2236"
author: "Ruiqi Xue"
date: "2023-10-11"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
```


## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```

```{r}
instacart = 
  instacart |> 
  as_tibble()
```


This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.


```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```


Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```


## Problem 2
Import data
```{r}
library(p8105.datasets)
data("brfss_smart2010")
```

Clean the dataset
```{r}
brfss_smart2010 =
  brfss_smart2010 |> 
  as_tibble() |>
  janitor::clean_names() |>
  rename("location_abbr" = "locationabbr") |>
  rename("location_desc" = "locationdesc")

brfss_oh = 
  filter(brfss_smart2010, topic == "Overall Health")
 
```

Check values of `response`
```{r}
print(unique(brfss_oh$response))
```

Keep cleaning the data
```{r}
brfss_oh = 
  filter(brfss_oh, response == "Excellent" |
           response == "Very good" |
           response == "Good" |
           response == "Fair" |
           response == "Poor")


response_vector = pull(brfss_oh, response)

response_factor = factor(response_vector, 
                         levels = c("Poor", "Fair", "Good", "Very good", "Excellent"),
                         ordered = TRUE)
brfss_oh$response = response_factor


```

Check which states were observed at 7 or more locations in 2002 and 2010.

```{r}
county_2002 = brfss_oh |>  
  filter(year == 2002) |> 
  group_by(location_abbr, location_desc) |> 
  summarise()
  

county_2002 = county_2002 |> 
  group_by(location_abbr) |> 
  summarise(count = n()) |> 
  filter(count >=7) 

print(county_2002)

```

The states that were observed at 7 or more locations in 2002 are: `r county_2002$location_abbr`.

```{r}
county_2010 = brfss_oh |>  
  filter(year == 2010) |> 
  group_by(location_abbr, location_desc) |> 
  summarise()
  

county_2010 = county_2010 |> 
  group_by(location_abbr) |> 
  summarise(count = n()) |> 
  filter(count >=7) 

print(county_2010)

```

Similarly, we get states that were observed at 7 or more locations in 2010: `r county_2010$location_abbr`.

Now make a "spaghetti” plot of average data value over time within a state.
```{r}
excellent_df = brfss_oh |> 
  filter(response == "Excellent") |> 
  select(location_abbr, year, data_value)

excellent_df = excellent_df |> 
  group_by(location_abbr, year) |> 
  summarise(mean_value = mean(data_value)) |> 
  rename(state = location_abbr)

excellent_df |> 
  ggplot(aes(x = year, y = mean_value, color = state)) +
  geom_line() + 
  theme(legend.position = "bottom") +
  labs(
    title = "'Spaghetti' Plot of Average Data Value Over Time for Each State",
    x = "Year",
    y = "Mean Data Value (of BHR)",
    color = "State"
  ) 

```

Now, make the two-panel plot showing for the years 2006 and 2010, the distribution of `data_value` for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
NY_response = 
  brfss_oh |> 
  filter(year == 2006 | year == 2010, location_abbr == "NY") |> 
  select(location_desc, response, data_value, year) 

NY_response |>
  ggplot(aes(x = response, y = data_value, color = location_desc)) + 
  geom_point() +
  theme(legend.position = "bottom") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  facet_grid(~year) +
  labs(
    title = "Distribution of Data Value for Responses Among Locations in NY State",
    x = "Response",
    y = "Data Value (of behavioral health risk)",
    color = "Location"
  ) 
```


## Problem 3

First, clean the datasets and merge them into one dataset.
```{r}
demographic_df = read_csv("data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female")) |>
  mutate(
    education = case_match(
      education,
      1 ~ "Less than high school",
      2 ~ "High school equivalent",
      3 ~ "More than high school"
    )) |>
  drop_na() |>
  filter(age >= 21)


education_factor = factor(pull(demographic_df, education), 
                          levels = c("Less than high school", "High school equivalent", "More than high school"), ordered = TRUE)

demographic_df$education = education_factor

accel_df = 
  read_csv("data/nhanes_accel.csv") |>
  janitor::clean_names() |>
  drop_na()

merged_df = left_join(demographic_df, accel_df)

merged_df2 = merged_df |>
  pivot_longer(
    min1:min1440,
    names_to = "minute",
    values_to = "mims_value"
  ) 
```




